---
title: "Machine Learning Modeling of Selected Weight-lifting Activites"
author: "Antonio Rubiera"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this analysis, we evaluate the data used by Velloso et al. in their paper "Qualitative Activity Recognition of Weight Lifting Exercises" [ACM SIGCHI 2013]. We will be using a specific seed (22) to ensure our data runs are reproducible. 

```{r library seed and data, warning=FALSE, message=FALSE}
library(tidyverse)
library(caret)
library(ggplot2)
library(knitr)
set.seed(22)
```

The training dataset has two types of data: a time series dataset, and time window measurements of this time series data. Here is the time series data.

```{r data time, warning=FALSE, message=FALSE}
training_all <- read.csv("./data/pml-training.csv")
training_all_data <- select(training_all, c("user_name",
                            "raw_timestamp_part_1","raw_timestamp_part_2",
                            "cvtd_timestamp","new_window","num_window","roll_belt",
                            "pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x",
                            "gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y",
                            "accel_belt_z","magnet_belt_x","magnet_belt_y",
                            "magnet_belt_z","roll_arm","pitch_arm","yaw_arm",
                            "total_accel_arm","gyros_arm_x","gyros_arm_y",
                            "gyros_arm_z","accel_arm_x","accel_arm_y","accel_arm_z",
                            "magnet_arm_x","magnet_arm_y","magnet_arm_z","roll_dumbbell",
                            "pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell",
                            "gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z",
                            "accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",                                                    "magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z",
                            "roll_forearm","pitch_forearm","yaw_forearm",
                            "total_accel_forearm","gyros_forearm_x",
                            "gyros_forearm_y","gyros_forearm_z","accel_forearm_x",
                            "accel_forearm_y","accel_forearm_z","magnet_forearm_x",
                            "magnet_forearm_y","magnet_forearm_z","classe"))
head(training_all_data)
str(training_all_data)
```

Here are the time window measurements after removal of several columns that only contained zeros or NAs:

--kurtosis_yaw_belt
--skewness_yaw_belt
--kurtosis_yaw_dumbbell
--skewness_yaw_dumbbell
--kurtosis_yaw_forearm
--skewness_yaw_forearm
--amplitude_yaw_dumbbell
--amplitude_yaw_forearm
--amplitude_yaw_belt

```{r data sdvar, warning=FALSE, message=FALSE}
training_all_clean <- read.csv("./data/pml-training.csv",
                               na.strings = c("NA","NaN","","#DIV/0!"))
training_nonblanks_clean <- filter(training_all_clean, kurtosis_roll_belt != "")
training_nonblanks_clean <- mutate(training_nonblanks_clean, 
                   rel_time = as.numeric(paste(raw_timestamp_part_1,
                   raw_timestamp_part_2,sep=".")))
training_sdvar_sel_clean_all <- select(training_nonblanks_clean,c("user_name",
        "classe","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp",
        "new_window","num_window","kurtosis_roll_belt","kurtosis_picth_belt",
        "skewness_roll_belt",
        "skewness_roll_belt.1","max_roll_belt","max_picth_belt",
        "max_yaw_belt","min_roll_belt","min_pitch_belt","min_yaw_belt","amplitude_roll_belt",
        "amplitude_pitch_belt","var_total_accel_belt","avg_roll_belt",
        "stddev_roll_belt","var_roll_belt","avg_pitch_belt","stddev_pitch_belt","var_pitch_belt",
        "avg_yaw_belt","stddev_yaw_belt","var_yaw_belt","var_accel_arm","avg_roll_arm",
        "stddev_roll_arm","var_roll_arm","avg_pitch_arm","stddev_pitch_arm","var_pitch_arm",
        "avg_yaw_arm","stddev_yaw_arm","var_yaw_arm","kurtosis_roll_arm","kurtosis_picth_arm",
        "kurtosis_yaw_arm","skewness_roll_arm","skewness_pitch_arm","skewness_yaw_arm",
        "max_roll_arm","max_picth_arm","max_yaw_arm","min_roll_arm","min_pitch_arm",
        "min_yaw_arm","amplitude_roll_arm","amplitude_pitch_arm","amplitude_yaw_arm",
        "kurtosis_roll_dumbbell","kurtosis_picth_dumbbell",
        "skewness_roll_dumbbell","skewness_pitch_dumbbell",
        "max_roll_dumbbell","max_picth_dumbbell","max_yaw_dumbbell","min_roll_dumbbell",
        "min_pitch_dumbbell","min_yaw_dumbbell","amplitude_roll_dumbbell","amplitude_pitch_dumbbell",
        "total_accel_dumbbell","var_accel_dumbbell","avg_roll_dumbbell",
        "stddev_roll_dumbbell","var_roll_dumbbell","avg_pitch_dumbbell","stddev_pitch_dumbbell",
        "var_pitch_dumbbell","avg_yaw_dumbbell","stddev_yaw_dumbbell","var_yaw_dumbbell",
        "kurtosis_roll_forearm","kurtosis_picth_forearm",
        "skewness_roll_forearm","skewness_pitch_forearm",
        "max_roll_forearm","max_picth_forearm","max_yaw_forearm","min_roll_forearm",
        "min_pitch_forearm","min_yaw_forearm","amplitude_roll_forearm","amplitude_pitch_forearm",
        "total_accel_forearm","var_accel_forearm","avg_roll_forearm",
        "stddev_roll_forearm","var_roll_forearm","avg_pitch_forearm","stddev_pitch_forearm",
        "var_pitch_forearm","avg_yaw_forearm","stddev_yaw_forearm","var_yaw_forearm",
        "rel_time"))
kable(table(training_sdvar_sel_clean_all$user_name, training_sdvar_sel_clean_all$classe))
```

